---
title: "p8105_hw2_rs4025"
author: "Rebecca Silva"
date: "10/4/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Questions: 
- best way to print var names? I did names()

## Problem 1 

```{r}
#read and clean data
data_trash = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = 1,
                          range = "A2:N38") %>%
  janitor::clean_names() %>%
  filter( dumpster != "NA") %>% 
  mutate( sports_balls = as.integer(
    round(sports_balls)))

data_precip2017 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = 4,
                          range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate( year = 2017)

data_precip2018 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = 3,
                          range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate( year = 2018)
```

```{r}
# combine 2017 and 2018 data  (join or bind?)
  
data_precip = 
  bind_rows(data_precip2017, data_precip2018) %>%
  janitor::clean_names() %>% 
  mutate( month = str_to_lower( month.name[month] ))
  
```

```{r}
# total precip in 2018
total_precip = 
  data_precip %>%
  filter( year == 2018) %>% 
  summarize( total = sum(total, na.rm = T))

# median number of sports balls in a dumpster in 2017
median = 
  data_trash %>% 
  filter( year == 2017) %>% 
  summarize( median = median(sports_balls, na.rm = T))
```


The data set named 'data_trash' gives information about trash disposal in specified dumpsters. The data set has `r nrow(data_trash)` observations and `r ncol(data_trash)` variables. Some key variables include the dumpster number, the date the trash is collected, and how much the trash weighs which are named `r names(data_trash)[1]`, `r names(data_trash)[4]`, and `r names(data_trash)[5]`, respectively. Many of the variables consist of frequent types of trash collected such as `r names(data_trash)[7]`, `r names(data_trash)[9]` and `r names(data_trash)[11]`.

The data set named 'data_precip' gives information about the precipitation in 2017 and 2018. There are `r nrow(data_precip)` observations and `r ncol(data_precip)` variables, which are `r names(data_precip)[1]`, `r names(data_precip)[2]` and `r names(data_precip)[3]`.

The total precipitation in 2018 is `r total_precip` inches. The median number of sports balls in a dumpster is `r median` since there is no observation of a dumpster in 2017 in the data set 'data_trash'. 


