---
title: "p8105_hw2_rs4025"
author: "Rebecca Silva"
date: "10/4/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```


## Problem 1 

```{r, reading/cleaning data}
# dumpster data 
data_trash = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                          sheet = "Mr. Trash Wheel",
                          range = "A2:N408") %>%
  janitor::clean_names() %>%
  filter( dumpster != "NA") %>% 
  mutate( sports_balls = as.integer(
    round(sports_balls)))

# precipitation 2017
data_precip2017 = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                          sheet = "2017 Precipitation",
                          range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate( year = 2017 )

# precipitation 2018
data_precip2018 = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                          sheet = "2018 Precipitation",
                          range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate( year = 2018 ) %>% 
  drop_na( total )

```

```{r}
# combine 2017 and 2018 data
data_precip = 
  bind_rows(data_precip2017, data_precip2018) %>%
  janitor::clean_names() %>% 
  mutate( month = str_to_lower( month.name[month] )) %>% 
  select( year, month, total)
```

```{r}
# total precip in 2018
total_precip = 
  data_precip %>%
  filter( year == 2018) %>% 
  summarize( total = sum(total, na.rm = T))

# median number of sports balls in a dumpster in 2017
median = 
  data_trash %>% 
  filter( year == 2017) %>% 
  summarize( median = median(sports_balls, na.rm = T))
```


The data set, `data_trash`, gives information about trash disposal in specified dumpsters. The data set has `r nrow(data_trash)` observations and `r ncol(data_trash)` variables. Some key variables include the dumpster number, the date the trash is collected, and how much the trash weighs which are named `r names(data_trash)[1]`, `r names(data_trash)[4]`, and `r names(data_trash)[5]`, respectively. Many of the variables consist of frequent types of trash collected such as `r names(data_trash)[7]`, `r names(data_trash)[9]` and `r names(data_trash)[11]`. The median number of sports balls in a dumpster is `r median`.

The data set, `data_precip`, gives information about the total precipitation in inches in 2017 and 2018 for each month. There are `r nrow(data_precip)` observations and `r ncol(data_precip)` variables, which are `r names(data_precip)[1]`, `r names(data_precip)[2]` and `r names(data_precip)[3]`. The total precipitation in 2018 is `r total_precip` inches.


## Problem 2

```{r, message= FALSE}
# pols data
pols = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, 
           into = c("year", "month", "day"), 
           sep = "-") %>% 
  mutate( month = str_to_lower(month.name[as.numeric(month)]),
          year = as.numeric(year),
          president = recode(prez_gop, 
                             `1` = "gop", 
                             `2` = "gop", 
                             `0` = "dem")) %>% 
  select(year, month, president, everything(),  - c(day, prez_gop, prez_dem)) 
```


```{r, message = FALSE}
# snp data
snp = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, 
           into = c("month", "day", "year"), 
           sep = "/") %>% 
  mutate( month = str_to_lower(month.name[as.numeric(month)]),
          year = as.numeric(year)) %>% 
  arrange( year, month) %>% 
  select( year, month, close)

```

```{r, message = FALSE}
# unemployment data 
unemployment = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  rename( january = jan, 
          febuary = feb, 
          march = mar,
          april = apr, 
          june = jun,
          july = jul,
          august = aug, 
          september = sep, 
          october = oct, 
          november = nov, 
          december = dec) %>% 
  pivot_longer(january:december, 
               names_to = "month", 
               values_to = "percent_unemployed") 
```


```{r}
# join data 
final_join = 
  left_join(
    left_join(pols, snp,      # first join 
              by = c("year", "month")), 
    unemployment, 
    by = c("year", "month"))

# order with key variables first
final_join = 
  final_join %>% 
  select(year, month, president:percent_unemployed, everything() )

```


The `pol` dataset was initially a dataset with 9 variables giving counts of democratic and republican politicians from the years `r min(pull(pols, year))` to `r max(pull(pols, year))`. The date column, labeled `mon`, became separate variables for year and month and we condensed two columns which indicate whether a president was a democrat or republican, labeling `president` as the variable that indicates their affiliate party. Note, there were 5 observations where `prez_gop` = 2 and `prez_dem` = 0, so I assume a '2' was meant to be '1', meaning the president was republican. The resulting dataset has `r nrow(pols)` observations of `r ncol(pols)` variables. 

The `snp` dataset initially contained `r nrow(snp)` observations of 2 variables: date, and closing value of S&P index (labeled `close`). We tidied the date column to contain variables `year` and `month` and arranged the observations by year and month. The dataset now contains `r ncol(snp)` variables and ranges from the years `r min(pull(snp, year))` to `r max(pull(snp, year))`, still with `r nrow(snp)` observations. 

Lastly, the `unemployment` data contained the unemployment percentage for 69 months in the years ranging from `r min(pull(unemployment, year))` to `r max(pull(unemployment, year))`. Each month was a separate column which we condensed into one named `month` and added the variable `percent_unemployed` to contain the values of each month variable. This resulted in more observations, specifically `r nrow(unemployment)` observations, and only `r ncol(unemployment)` variables. 

We merged all three of the above datasets into the dataset `final_join`. The final dataset contains `r nrow(final_join)` observations of `r ncol(final_join)` variables which includes observations from the year `r min(pull(final_join, year))` to `r max(pull(unemployment, year))`. We are most interested in the variables `year`, `month`, `president`, `logical`, `close`, and `percent_unemployed`, all previously described. Due to the nature of joining data, we are left with some NA's, specifically `r sum(is.na(final_join))` NAs, most coming from specific years when that data might not have been collected. 

## Problem 3 

```{r}
# read in dataset 
baby_names = 
  read_csv("./data/Popular_Baby_Names.csv", 
           col_types = "dcccdd") %>% 
  janitor::clean_names() %>%
  mutate( gender = str_to_lower(gender),
          childs_first_name = str_to_lower(childs_first_name),
          ethnicity = str_to_lower(ethnicity),
          ethnicity = replace( ethnicity, ethnicity == "white non hisp", "white non hispanic"), 
          ethnicity  = replace( ethnicity, ethnicity == "black non hisp", "black non hispanic"), 
          ethnicity  = replace( ethnicity, ethnicity == "asian and paci", "asian and pacific islander")) %>%
  distinct(.keep_all = TRUE) # remove duplicate rows 
```


```{r}
# dataset for rank of name Olivia 
olivia = 
  baby_names %>% 
  filter(childs_first_name == "olivia", gender == "female") %>% 
  mutate(ethnicity = str_to_title(ethnicity)) %>%
  rename(Ethnicity = ethnicity) %>% 
  select(-c(count, childs_first_name, gender)) %>% 
  pivot_wider(
    names_from = "year_of_birth", 
    values_from = "rank"
  )
  
knitr::kable(olivia, 
             caption = "Table 1: Rank in Popularity of 'Olivia' Over Time")
```

```{r}
# dataset for most popular male names 
boy_names = 
  baby_names %>% 
  filter( rank == 1, gender == "male") %>% 
  mutate( childs_first_name = str_to_title(childs_first_name), 
          ethnicity = str_to_title(ethnicity)) %>%
  rename(Ethnicity = ethnicity) %>% 
  group_by(year_of_birth, Ethnicity) %>% 
  select( - c(gender, count, rank)) %>% 
  pivot_wider(
    names_from = "year_of_birth", 
    values_from = "childs_first_name"
  )

knitr::kable(boy_names, 
             caption = "Table 2: Most Popular Boy Names Over Time" ) 
```


```{r, message = FALSE }
# tidy dataset for plot 
plot_data = 
  baby_names %>% 
  filter( gender == "male", 
          ethnicity == "white non hispanic",
          year_of_birth == 2016)

# scatter plot 
ggplot(plot_data, aes(x = rank, y = count)) + 
  geom_point() + 
  labs(
    title = "Frequency vs Rank of White Non-hispanic Boy Names in 2016",
    x = "Rank of Name",
    y = "Frequncy of Name"
  ) +
 theme(plot.title = element_text(hjust = .5))
```

