---
title: "p8105_hw2_rs4025"
author: "Rebecca Silva"
date: "10/4/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Questions: 
- best way to print var names? I did names()
- do we have to specify col_types or can we let R do that 
- convert month abrrev to full month 
- make year as.numeric?

## Problem 1 

```{r}
#read and clean data
data_trash = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = 1,
                          range = "A2:N38") %>%
  janitor::clean_names() %>%
  filter( dumpster != "NA") %>% 
  mutate( sports_balls = as.integer(
    round(sports_balls)))

data_precip2017 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = 4,
                          range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate( year = 2017)

data_precip2018 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = 3,
                          range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate( year = 2018)
```

```{r}
# combine 2017 and 2018 data  (join or bind?)
  
data_precip = 
  bind_rows(data_precip2017, data_precip2018) %>%
  janitor::clean_names() %>% 
  mutate( month = str_to_lower( month.name[month] ))
  
```

```{r}
# total precip in 2018
total_precip = 
  data_precip %>%
  filter( year == 2018) %>% 
  summarize( total = sum(total, na.rm = T))

# median number of sports balls in a dumpster in 2017
median = 
  data_trash %>% 
  filter( year == 2017) %>% 
  summarize( median = median(sports_balls, na.rm = T))
```


The data set named 'data_trash' gives information about trash disposal in specified dumpsters. The data set has `r nrow(data_trash)` observations and `r ncol(data_trash)` variables. Some key variables include the dumpster number, the date the trash is collected, and how much the trash weighs which are named `r names(data_trash)[1]`, `r names(data_trash)[4]`, and `r names(data_trash)[5]`, respectively. Many of the variables consist of frequent types of trash collected such as `r names(data_trash)[7]`, `r names(data_trash)[9]` and `r names(data_trash)[11]`.

The data set named 'data_precip' gives information about the precipitation in 2017 and 2018. There are `r nrow(data_precip)` observations and `r ncol(data_precip)` variables, which are `r names(data_precip)[1]`, `r names(data_precip)[2]` and `r names(data_precip)[3]`.

The total precipitation in 2018 is `r total_precip` inches. The median number of sports balls in a dumpster is `r median` since there is no observation of a dumpster in 2017 in the data set 'data_trash'. 

## Problem 2

```{r}
# pols data
pols = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, 
           into = c("year", "month", "day"), 
           sep = "-") %>% 
  mutate( month = str_to_lower(month.name[as.numeric(month)]),
          year = as.numeric(year)) %>% 
  pivot_longer(c(prez_gop, prez_dem), 
               names_to = "president", 
               names_prefix = "prez_", 
               values_to = "logical") %>% #what to name?
  select( - day)

```


```{r}
# snp data
snp = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, 
           into = c("month", "day", "year"), 
           sep = "/") %>% 
  mutate( month = str_to_lower(month.name[as.numeric(month)]),
          year = as.numeric(year)) %>% 
  arrange( year, month) %>% 
  select( year, month, close)

```

```{r}
# unemployment data 
unemployment = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  rename( january = jan, 
          febuary = feb, 
          march = mar,
          april = apr, 
          june = jun,
          july = jul,
          august = aug, 
          september = sep, 
          october = oct, 
          november = nov, 
          december = dec) %>% 
  pivot_longer(january:december, 
               names_to = "month", 
               values_to = "percent_unemployed") 
```


```{r}
# join data 

final_join = 
  left_join(
    left_join(pols, snp,  # first join 
              by = c("year", "month")), 
    unemployment, 
    by = c("year", "month"))

# order with key variables first
final_join = 
  final_join %>% 
  select(year, month, president:percent_unemployed, everything() )

```

 Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

The 'pol' dataset was initially a dataset with 9 variables giving counts of democratic and republican politicians from `r min(pull(pols, year))` to `r max(pull(pols, year))`. The 'date' column labeled 'mon' became separate variables for year and month and we condensed two columns which indicate whether a president was a democrat or republican, labeling 'president' as the variable that indicates their affiliate party and 'logical' as the variable representing a '1' for democrat and '0' for republican. The resulting dataset has `r nrow(pols)` observations of `r ncol(pols)` variables. 

The 'snp' data initially contained `r nrow(snp)` observations of two variables: date, and closing value of S&P index (labeled 'close'). We tidied the date column to contain variables 'year' and 'month' and arranged the observations by year and month. The dataset now contains `r ncol(snp)` variables and ranges from `r min(pull(snp, year))` to `r max(pull(snp, year))`. 

Lastly, the 'unemployment' data contained the unemployment percentage for 69 specified months of years ranging from `r min(pull(unemployment, year))` to `r max(pull(unemployment, year))`. Each month was a separate column which we condensed into one named 'month' and added the variable 'percent_unemployed' to contain the values of each month variable. This resulted in more observations, specifically `r nrow(unemployment)` observations, and only `r ncol(unemployment)` variables. 

We merged all three of the above datasets into the dataset 'final_join'. The final dataset contains `r nrow(final_join)` observations of `r ncol(final_join)` variables which includes observations from `r min(pull(final_join, year))` to `r max(pull(unemployment, year))`. We are most interested in the variables 'year', 'month', 'president', 'logical', 'close', and 'percent_unemployed', all previously described. Due to the nature of joining data, we are left with NA's, most coming from specific years when that data might not have been collected. 


